\section{Data Description}
\subsection{Source of Data}
The dataset used for this project comprises approximately 17,900 labeled images of crops and their associated diseases. These images are sourced from \href{https://www.kaggle.com/datasets/mexwell/crop-diseases-classification}{Kaggle}. Additionally, the dataset may be supplemented with synthetic data generated through data augmentation techniques to improve model performance and generalization.



\subsection{Type of Data}
The dataset primarily consists of unstructured image data. Each image belongs to one of five distinct classes, representing specific crop diseases. The data is organized in the following format:
\begin{itemize}
    \item \textbf{Input Data}: High-resolution images of crops captured under various environmental conditions.
    \item \textbf{Labels}: Each image is annotated with its corresponding crop disease class, enabling supervised learning.
\end{itemize}

\subsection{Data Preprocessing Requirements}
To ensure optimal model performance and reliability, the dataset will undergo the following preprocessing steps:
\begin{itemize}
    \item \textbf{Image Resizing}: All images will be resized to a uniform dimension to standardize input for the machine learning model.
    \item \textbf{Normalization}: Pixel values will be normalized to a common scale to enhance model convergence during training.
    \item \textbf{Data Augmentation}: Techniques such as rotation, flipping, and cropping will be applied to artificially increase dataset diversity and address class imbalance.
    \item \textbf{Noise Removal}: Images with poor quality, excessive noise, or irrelevant content will be filtered out.
    \item \textbf{Splitting}: The dataset will be divided into training, validation, and test sets to evaluate model performance effectively.
\end{itemize}

By ensuring proper preprocessing and leveraging diverse image data, the project aims to build a robust and accurate crop disease detection system.
