\section{Implementation Plan}
This project will be executed in several phases, each focusing on a critical aspect of the crop disease prediction system. The steps and their respective timelines are as follows:

\subsection{Phases of Implementation}
\begin{enumerate}
    \item \textbf{Data Collection and Understanding (Week 1–2)}:
    \begin{itemize}
        \item Collect publicly available datasets containing images of crop diseases.
        \item Analyze the dataset structure, distribution, and class balance.
        \item Supplement the dataset with synthetic data using data augmentation techniques if required.
    \end{itemize}

    \item \textbf{Data Preprocessing (Week 3–4)}:
    \begin{itemize}
        \item Standardize image dimensions by resizing.
        \item Normalize pixel values for consistent input across the model.
        \item Remove noisy or irrelevant data and handle class imbalance using augmentation.
        \item Split the dataset into training (70\%), validation (15\%), and test (15\%) subsets.
    \end{itemize}

    \item \textbf{Model Selection and Training (Week 5–8)}:
    \begin{itemize}
        \item Evaluate various machine learning models, focusing on convolutional neural networks (CNNs).
        \item Train selected models on the preprocessed dataset using GPUs for faster computation.
        \item Fine-tune hyperparameters, such as learning rate, batch size, and number of epochs, to optimize performance.
    \end{itemize}

    \item \textbf{Model Testing and Evaluation (Week 9–10)}:
    \begin{itemize}
        \item Evaluate the trained model on the test dataset using metrics such as accuracy, precision, recall, and F1 score.
        \item Conduct error analysis to identify common misclassifications and refine the model.
    \end{itemize}

    \item \textbf{System Deployment (Week 11–12)}:
    \begin{itemize}
        \item Develop a user-friendly interface for farmers and agricultural stakeholders.
        \item Integrate the trained model into the system for real-time crop disease detection.
        \item Test the deployed system in a simulated or real-world environment to ensure reliability.
    \end{itemize}
\end{enumerate}
\newpage
\subsection{Estimated Timeline}
\begin{longtable}{|c|p{4cm}|c|c|c|p{4cm}|}
\hline
\textbf{ID} & \textbf{Name} & \textbf{Start Date} & \textbf{Finish Date} & \textbf{Duration} & \textbf{Notes} \\ \hline
1 & Define objectives, scope, and deliverables & 2024-12-02 & 2024-12-09 & 6 days & Identify goals and project boundaries. \\ \hline
2 & Gather image datasets requirements & 2024-12-10 & 2024-12-17 & 6 days & Determine data specifications and needs. \\ \hline
3 & Finalize workflow and timeline & 2024-12-18 & 2024-12-25 & 6 days & Confirm task sequence and deadlines. \\ \hline
4 & Collect datasets & 2024-12-23 & 2025-01-03 & 10 days & Acquire all necessary datasets. \\ \hline
5 & Preprocess data & 2025-01-02 & 2025-01-08 & 5 days & Clean and prepare datasets for training. \\ \hline
6 & Split datasets & 2025-01-09 & 2025-01-15 & 5 days & Divide data into training and testing sets. \\ \hline
7 & Research/select image classification model & 2025-01-16 & 2025-01-22 & 5 days & Evaluate and choose a suitable model. \\ \hline
8 & Design/implement architecture & 2025-01-16 & 2025-01-22 & 5 days & Develop the system architecture. \\ \hline
9 & Split the test data and train data & 2025-01-23 & 2025-01-28 & 4 days & Create separate datasets for testing and training. \\ \hline
10 & Train the model & 2025-01-29 & 2025-02-11 & 10 days & Train the model using the prepared datasets. \\ \hline
11 & Tune hyperparameters & 2025-02-12 & 2025-02-18 & 5 days & Optimize model performance. \\ \hline
12 & Evaluate the model & 2025-02-19 & 2025-02-25 & 5 days & Assess model accuracy and performance. \\ \hline
13 & Fine-tune the model & 2025-02-19 & 2025-02-25 & 5 days & Improve model based on evaluation results. \\ \hline
14 & Test the model & 2025-02-26 & 2025-03-04 & 5 days & Validate model using test data. \\ \hline
15 & Analyze/report metrics & 2025-03-05 & 2025-03-11 & 5 days & Document performance metrics. \\ \hline
16 & Deploy the model & 2025-03-12 & 2025-03-18 & 5 days & Deploy the final model in production. \\ \hline
17 & Integrate/test real-time inputs & 2025-03-12 & 2025-03-18 & 5 days & Test the model with live data. \\ \hline
18 & Prepare documentation/report & 2025-03-19 & 2025-03-25 & 5 days & Create technical and user documentation. \\ \hline
19 & Create presentation & 2025-03-19 & 2025-03-25 & 5 days & Prepare slides for project presentation. \\ \hline
20 & Develop an application & 2025-03-26 & 2025-04-01 & 5 days & Build an application to integrate the model. \\ \hline
\end{longtable}


By following this structured implementation plan, the project aims to achieve a robust and efficient solution for crop disease prediction within the specified timeline.
